---
title: "Introduction to mlr3"
output:
  html_document:
    toc: TRUE
---

```{r, include = FALSE, warning = FALSE, message = FALSE}
# Just some preparation
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>"
)
if (require("data.table")) data.table::setDTthreads(1)
options(width = 90)
set.seed(8008135)
lgr::get_logger("mlr3")$set_threshold("warn")

# Create the train and test data
# task = OpenML::getOMLTask(75232)
# task2 = OpenML::convertOMLTaskToMlr(task)
# train_set = task2$mlr.rin$train.inds[[1]]
# test_set = task2$mlr.rin$test.inds[[1]]
# d = mlr::getTaskData(task2$mlr.task)
# d$Class = factor(d$Class, levels = c("2","1"))
# train = d[train_set, ]
# test = d[test_set, ]
# saveRDS(train, "qsar_train.rds")
# saveRDS(test, "qsar_test.rds")

train = readRDS("qsar_train.rds")
test = readRDS("qsar_test.rds")
solution = F
set.seed(1)
library(mlr3)
```

## Exercise 1

Read the description of the data at https://archive.ics.uci.edu/ml/datasets/QSAR+biodegradation and get an overview of the data by looking at some simple univariate visualizations of the features at the webpage https://www.openml.org/d/1494 (open the link with chrome and click on *'show all 42 features'* in the features section).

## Exercise 2
The whole data consists of `r nrow(train) + nrow(test)` observations (molecules). 
We have split the data into training data `qsar_train.rds` and test data `qsar_test.rds`.
For now, consider only the training data consisting of `r nrow(train)` molecules, each of it belonging either to the experimental class "not ready biodegradable" or to the class "ready biodegradable".
The ultimate goal is to predict the `Class` column with values 1 (not ready biodegradable) and 2 (ready biodegradable) using information on molecular descriptors (i.e., our `r ncol(train)-1` features).

a. Import the training data into R using the code below and create a classification task using `mlr3`:
```{r, eval = solution, echo = solution, message=FALSE}
# Importing the train data
train = readRDS("qsar_train.rds")
```
```{r, eval = solution, echo = solution}
library(mlr3)
task = TaskClassif$new("qsar", train, "Class")
```
b. Define a random forest learner, which will be able to **predict the probability** of a molecule belonging one of the two classes. (Note: don't forget to use the `mlr3learners` package)
```{r, eval = solution, echo = solution}
# if you don't load mlr3learners, you won't find a random forest
mlr_learners

# after loading mlr3learners, the available list of learners is extended
library(mlr3learners)
mlr_learners

# now, we can construct the random forest learner
learner_rf = lrn("classif.ranger", predict_type = "prob")
```
c. Before we can assess the performance of the learner, we need to choose one or more performance measures we are interested in.
    - First, look up the names (i.e., `Keys`) of all implemented performance measures:
      ```{r}
mlr_measures
```
      Try to find the "classification accuracy" and the "area under the roc curve" measure from this list by looking at the help page of the corresponding measure using `?mlr_measures_xxx` where `xxx` is the name of the measure.
    - Now, define a list containing these two performance measures
      ```{r, eval = solution, echo = solution}
msr_list = list(msr("classif.auc"), msr("classif.acc"))
```
d. Use 10-fold cross-validation to estimate the performance of your random forest on the training data.
```{r, eval = solution, echo = solution}
# define the resampling strategy
resampling = rsmp("cv", folds = 10)
# run the resampling
res_sub = resample(task, learner_rf, resampling)
# calculate the performance of the two measures from the previous exercise
res_sub$aggregate(msr_list)
```

e. Sofar, we have only used the available training data to estimate the performace of a random forest. Now, do the following steps:
    - Train the random forest on the whole training data.
    - Import the test data and predict the probabilities of belonging to one of the two classes for the molecules in the test data.
    - Measure the performance of the trained random forest on the test data using the "classification accuracy" and the "area under the roc curve".
```{r, eval = solution, echo = solution}
# currently we havn't trained a model
learner_rf$model

# we use the training data (included in task) to train a model
learner_rf$train(task)
learner_rf$model

# Import the test data
test = readRDS("qsar_test.rds")
# we make predictions for the test data
preds = learner_rf$predict_newdata(test)
preds

# measure the performance on the test data
preds$score(msr_list)
```

## Exercise 3

Compare 2 different learners using the `benchmark` function.

```{r, eval = solution, message=FALSE, warning=FALSE, echo = solution}
lrn_list = list(
  lrn("classif.log_reg", predict_type = "prob"),
  lrn("classif.ranger", predict_type = "prob")
)

bm_design = benchmark_grid(task = task, resamplings = resampling, learners = lrn_list)
bmr = benchmark(bm_design)
bmr$aggregate(msr_list)
```
