---
  title: "Introduction to mlr3"
output:
  html_document:
  toc: TRUE
---

  ```{r, include = FALSE, warning = FALSE, message = FALSE}
library(mlr3)
train = readRDS("qsar_train.rds")
test = readRDS("qsar_test.rds")
```

## Exercise 1

Read the description of the data at https://archive.ics.uci.edu/ml/datasets/QSAR+biodegradation and get an overview of the data by looking at some simple univariate visualizations of the features at the webpage https://www.openml.org/d/1494 (open the link with chrome and click on *'show all 42 features'* in the features section).

## Exercise 2
The whole data consists of `r nrow(train) + nrow(test)` observations (molecules).
We have split the data into training data `qsar_train.rds` and test data `qsar_test.rds`.
For now, consider only the training data consisting of `r nrow(train)` molecules, each of it belonging either to the experimental class "not ready biodegradable" or to the class "ready biodegradable".
The ultimate goal is to predict the `Class` column with values 1 (not ready biodegradable) and 2 (ready biodegradable) using information on molecular descriptors (i.e., our `r ncol(train)-1` features).

a. Import the training data into R using the code below and create a classification task using `mlr3`:
  ```{r, message=FALSE}
# Import data
# Create task
```
b. Define a random forest learner, which will be able to **predict the probability** of a molecule belonging one of the two classes. (Note: don't forget to use the `mlr3learners` package)
```{r}
# Define learner
```
c. Before we can assess the performance of the learner, we need to choose one or more performance measures we are interested in.
    - First, look up the names (i.e., `Keys`) of all implemented performance measures:
```{r}
mlr_measures
```
      Try to find the "classification accuracy" and the "area under the roc curve" measure from this list by looking at the help page of the corresponding measure using `?mlr_measures_xxx` where `xxx` is the name of the measure.
    - Now, define a list containing these two performance measures
```{r}
# Create list of measures
```
d. Use 10-fold cross-validation to estimate the performance of your random forest on the training data.
```{r}
# Define the resampling strategy and run the resampling
```

e. Sofar, we have only used the available training data to estimate the performace of a random forest. Now, do the following steps:
    - Train the random forest on the whole training data.
    - Import the test data and predict the probabilities of belonging to one of the two classes for the molecules in the test data.
    - Measure the performance of the trained random forest on the test data using the "classification accuracy" and the "area under the roc curve".
```{r}
# 1. Train random forest on training data
# 2. Import test data
# 3. Predict probabilities for test data
# 4. Measure performance on test data
```

## Exercise 3

Compare 2 different learners using the `benchmark` function.

```{r}
# Benchmark two learners
```
